\pkg{RProtoBuf}: Efficient Cross-Language Data Serialization in \proglang{R}

### 2014-01-23 papers/jss/article.Rnw
Add new, better ack for saptarshi.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@839 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
Switch around a sentence to get a better flow
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@840 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

A more modern, widely used format is \emph{JavaScript Object
  Notation} (\texttt{JSON}), which is derived from the object literals of
\proglang{JavaScript}, and used increasingly on the world wide web. \texttt{JSON} natively
supports arrays and distinguishes 4 primitive types: numbers, strings,
booleans, and null. However, as it too is a text-based format, numbers are
stored as human-readable decimal notation which is inefficient and
leads to loss of type (double versus integer) and precision. Several \proglang{R} packages
implement functions to parse and generate \texttt{JSON} data from \proglang{R}
objects \citep{rjson,RJSONIO,jsonlite}.
A number of binary formats based on \texttt{JSON} have been proposed
that reduce the parsing cost and improve efficiency, but these formats
are not widely supported.  Furthermore, such formats lack a separate
schema for the serialized data and thus still duplicate field names
with each message sent over the network or stored in a file.
% and still must send duplicate field names
% with each message since there is no separate schema.
% \pkg{MessagePack}
% and \pkg{BSON} both have \proglang{R}
% interfaces \citep{msgpackR,rmongodb}, but these formats lack a separate schema for the serialized
% data and thus still duplicate field names with each message sent over
% the network or stored in a file.  Such formats also lack support for
% versioning when data storage needs evolve over time, or when
% application logic and requirement changes dictate updates to the
%message format.

@@

The first versions of \CRANpkg{RProtoBuf} were written during 2009-2010.
Very significant contributions, both in code and design, were made by
Romain Fran\c{c}ois whose continued influence on design and code is
greatly appreciated. Several features of the package reflect
the design of the \CRANpkg{rJava} package by Simon Urbanek
The user-defined table mechanism, implemented by Duncan Temple Lang for the
purpose of the \pkg{RObjectTables} package, allows for the dynamic symbol lookup.
Kenton Varda was generous with his time in reviewing code and explaining
obscure protocol buffer semantics.  Karl Millar was very
helpful in reviewing code and offering suggestions.  
%The contemporaneous work by Saptarshi Guha on \pkg{RHIPE} was a strong
%initial motivator.

>>>>

A more modern, widely used format is \emph{JavaScript ObjectNotation} 
(\texttt{JSON}), which is derived from the object literals of
\proglang{JavaScript}, and used increasingly on the world wide web. 
Several \proglang{R} packages implement functions to parse and generate
\texttt{JSON} data from \proglang{R} objects \citep{rjson,RJSONIO,jsonlite}.
\texttt{JSON} natively supports arrays and 4 primitive types: numbers, strings,
booleans, and null. However, as it too is a text-based format, numbers are
stored as human-readable decimal notation which is inefficient and
leads to loss of type (double versus integer) and precision. 
A number of binary formats based on \texttt{JSON} have been proposed
that reduce the parsing cost and improve efficiency, but these formats
are not widely supported.  Furthermore, such formats lack a separate
schema for the serialized data and thus still duplicate field names
with each message sent over the network or stored in a file.
% and still must send duplicate field names
% with each message since there is no separate schema.
% \pkg{MessagePack}
% and \pkg{BSON} both have \proglang{R}
% interfaces \citep{msgpackR,rmongodb}, but these formats lack a separate schema for the serialized
% data and thus still duplicate field names with each message sent over
% the network or stored in a file.  Such formats also lack support for
% versioning when data storage needs evolve over time, or when
% application logic and requirement changes dictate updates to the
%message format.

@@

The first versions of \CRANpkg{RProtoBuf} were written during 2009-2010.
Very significant contributions, both in code and design, were made by
Romain Fran\c{c}ois whose continued influence on design and code is
greatly appreciated. Several features of the package reflect
the design of the \CRANpkg{rJava} package by Simon Urbanek
The user-defined table mechanism, implemented by Duncan Temple Lang for the
purpose of the \pkg{RObjectTables} package, allows for the dynamic symbol lookup.
Kenton Varda was generous with his time in reviewing code and explaining
obscure protocol buffer semantics.  Karl Millar was very
helpful in reviewing code and offering suggestions.  Saptarshi Guha's
work on RHIPE and implementation of a universal message type for \proglang{R}
language objects allowed us to add the \code{serialize_pb} and \code{unserialize\_pb}
methods for turning arbitrary R objects into protocol buffers without
a specialized pre-defined schema.
